Dysregulation of Cerebral Hemoglobin Content in Malawian Children with Malaria

### 2021-05-13 figures.Rmd
fig 1 rearrange and add KW p-values

library(tidyverse)
library(janitor)
library(factoextra)
library(ggrepel)
library(ggpubr)
library(tidymodels)
library(vip)
library(scales)
library(reshape)

@@

df_hbox_voi <- df_hbox_impute %>% 
  dplyr::select(c(Status, DeoxyHb, DeoxyHb_alpha, OxyHb, OxyHb_alpha, O2sat, O2sat_alpha, THC, THC_alpha, THC_sd)) %>% 
  dplyr::filter(DeoxyHb_alpha < 3) %>% 
  dplyr::rename("THC sd" = "THC_sd", "THCα" = "THC_alpha", "Deoxy Hb" = "DeoxyHb", "Oxy Hb" = "OxyHb", 
                "O2 sat" = "O2sat", "O2 sat α" = "O2sat_alpha", "Deoxy Hb α" = "DeoxyHb_alpha", "Oxy Hb α" = "OxyHb_alpha")

df_hbox_voi_long <- df_hbox_voi %>% 
  pivot_longer(2:10, names_to = "variable", values_to = "value") %>%
  mutate(variable = factor(variable, levels = c("Deoxy Hb", "Oxy Hb", "O2 sat", "THC", "THC sd", 
                                                "Deoxy Hb α", "Oxy Hb α", "O2 sat α", "THCα")))

@@

  ggplot(aes(x = Status, y = value)) +
  geom_violin(aes(color = Status)) +
  geom_jitter(position = position_jitter(0.2), shape = 1) +
  stat_summary(fun = "median", geom = "crossbar", aes(color = Status), size = 0.2, width = 0.5) +
  facet_wrap(~variable, scales = "free_y") +

@@

```
**Figure 1.** Differences in variables measured by Oxiplex and associated alpha values among patient groups. Note: removed one outlier from DeoxyHb alpha (3.52)

>>>>

library(tidyverse)
library(janitor)
library(FSA)
library(factoextra)
library(ggrepel)
library(ggpubr)
library(tidymodels)
library(vip)
library(scales)
library(reshape)

@@

df_hbox_voi <- df_hbox_impute %>% 
  dplyr::select(c(Status, `Deoxy Hb`, `Deoxy Hb α`, `Oxy Hb`, `Oxy Hb α`, `Cerebral O2 sat`, `Cerebral O2 sat α`, THC, `THCα`)) %>% 
  dplyr::filter(`Deoxy Hb α` < 3)

df_hbox_voi_long <- df_hbox_voi %>% 
  pivot_longer(2:9, names_to = "variable", values_to = "value") %>%
  mutate(variable = factor(variable, levels = c("Oxy Hb", "Oxy Hb α",   
                                                 "Deoxy Hb", "Deoxy Hb α", 
                                                "Cerebral O2 sat", "Cerebral O2 sat α", 
                                                "THC", "THCα")))

@@

  ggplot(aes(x = Status, y = value)) +
  geom_violin(aes(color = Status)) +
  geom_jitter(position = position_jitter(0.2), shape = 1) +
  stat_summary(fun = "median", geom = "crossbar", aes(color = Status), size = 0.2, width = 0.5) +
  facet_wrap(~variable, scales = "free_y", nrow = 4, ncol = 2) +

@@

```{r fig1.pvals, eval = FALSE}

# THC
kruskal.test(THC ~ Status, data = df_hbox_voi)
dunnTest(THC ~ Status, data = df_hbox_voi, method = "bh")

# Cerebral O2 sat
kruskal.test(`Cerebral O2 sat` ~ Status, data = df_hbox_voi)

**Figure 1.** Differences in variables measured by Oxiplex and associated α values among patient groups. THC levels were different between all three groups (Kruskal-Wallis adjusted p-value = 3.25x10^-6^; Dunn Test post-hoc adjusted p-values: CM vs HV = 1.66x10^-6^, CM vs UM = .032, HV vs UM = .033). Levels of cerebral O~2~ saturation did not differ among groups (Kruskal-Wallis adjusted p-value = .07). Note: removed one outlier from Deoxy Hb α (3.52)
